Toward theories of deliberated preferences

### 2020-11-01 simple.tex
Better Descr. th.

This article defines decision theories, called theories of \ac{DJ}, that are both normative and descriptive: they claim to be able to yield valid prescriptions, and to allow confrontation to empirical data.

@@

\Cref{sec:approaches} defines more precisely what is meant here with normative and descriptive theories. These definitions permit to readily see why the apparent impossibility of a claim being both normative and descriptive is an illusion, thereby enabling a solution to what could be called the “normative against descriptive” gap. It also reviews related works in the literature, and briefly sketches the current proposal.
\Cref{sec:normative,sec:descriptive} define the theories of \ac{DJ} and prove that these theories are normative and descriptive in the sense defined in \cref{sec:approaches}. \Cref{sec:preferences} define theories of \ac{DP} on the basis of theories of \ac{DJ}, and \cref{sec:conclusion} concludes.

@@

This gap can thus be overcome by considering reflective descriptive theories, instead of shallow descriptive theories. This is the path this article adopts.
It defines theories that are both normative and descriptive: they attempt to deliver valid prescriptions, and these prescriptions are claimed to predict reflective behavior.

@@

\section{Descriptive theories}
\label{sec:descriptive}
\subsection{Empirical theories, informally}
\commentOCf{TODO : intégrer les explications dans la partie formelle et les reformuler.}
A normative theory defines how to determine that an argument is not decisive for a given proposition: suffices to find a sequence of arguments that does not lead to that proposition \commentYM{pas compris : tu veux dire "une séquence qui se termine par cet argument et qui n'amène pas à cette proposition ?}. This, in turn, under condition of accepting its normative validity \commentYM{mais c'est quoi cette "normative validiy" ? de quelles conditions parles-tu ?}, permits to know something about the \ac{DJ} of an individual. But this gives no concrete means to know someone’s \ac{DJ}. The set of potentially relevant arguments will generally be very large, and can’t be explored thoroughfully \commentYM{certain que ce terme existe ?}. More importantly, only some sequences of arguments can be tested for a given individual. That is because, first, the individual’s patience is limited, and second, once a sequence has been presented to the individual, this cannot necessarily be undone: she might remember some of the arguments, and they may have a lasting influence on her perspective. Postulating that the individual forgets after some time and separating the tests in time might mitigate this problem, but not necessarily cancel it. \commentYM{je comprends ce que tu veux dire ici car ça fait un bout de temps qu'on parle de ces question. Je doute cependant qu'un lecteur innocent comprenne un traitre mot}

%As a result, it should be possible to determine someone’s \ac{DJ} using only a small increasing set of sequences of arguments. A set of sequences of arguments is increasing iff the sequences can be ordered completely such that each sequence following another one “continues” the other one: its first arguments are those constituting the sequence that it follows.

These two problems are tackled by introducing empirical theories of \ac{DJ}. An empirical theory for a given normative theory defines a subset of propositions of the agenda, that it claims are contained in the \acp{DJ} of individuals concerned by the normative theory. They are called the proposition that the theory supports. \commentYM{dans la mesure où, à ce stade, je ne comprends pas bien ce que contient une "normative theory", la séparation avec l' "empirical theory" n'est pas claire pour moi} The theory also claims that it is able to argue in favor of the proposition it supports in such a way that individuals will be convinced of the quality of these arguments. This is accomplished by three binary relations, part of the definition of empiricial theories: one represents its initial arguments, one plans the possible attacks, and one defines its counter-arguments. (An empirical theory also defines a concatenation operation able to dynamically build arguments from fragments of arguments, an aspect which is not discussed in this informal presentation.)
The theory claims that when one of its arguments in support of one of its claims is unconvincing to an individual, it is because of one of the anticipated attacks; and the theory claims to be able to provide counter-arguments to such attacks as they come, that will in turn be either convincing to the individual, or fail to be because of some anticipated attack. (The theory is also constrained to have “finite depth”, meaning that it must set an a priori limit to the number of such iterations of attacks and counter-arguments.)

Thanks to these operators, an empirical theory may adjust its argumentation to each individual \commentYM{je ne comprends pas cette phrase, et en particulier le sens qu'il faut donner au terme "adjust"}. Instead of having to give a single decisive argument in support of its claims, the theory may react to exactly the counter-arguments that a given individual thinks relevant. This task is easier, but still permits to obtain what is desired in this article: there exists a purely empirical property \commentYM{kezako ?}of a theory, called convincingness, that, together with an axiom about the normative adequacy of the observation protocol, implies the correctness of the claims of the theory. In more details, first, the claim of the theory, of being able to argue convincingly, can be checked by ensuring that no falsification instances exist, thus, it is a fully empirical claim (in the sense that it requires no axiom). Second, granted normative adequacy, if indeed a theory is convincing, then it is valid, meaning that the proposition that it claims are \acp{DJ} are indeed \acp{DJ}. These claims are proven in \cref{th:convnofals,th:valid}.

@@

\subsection{Falsification}
\label{sec:fals}
Given $\args \subseteq \allargs$ and $\hist \in \allhist$,
the set $\args \cap \range(\hist)$ denotes the arguments that are both in $\args$ and contained in the sequence $\hist$.

Define a falsification sequence for $i, \ar_0, \phi$, with $\ar_0 \gleadsto \phi$, thus $\ar_0$ an initial argument for $\phi$, as a finite sequence $\hist^1, \hist^3, …, \hist^{k + 1}$, $k \in 2\N$, such that it presents successive challenges, which means the following. 
Given $l \in 2\N \suchthat l ≤ k$, $\ar_l \in \allargs$ and $\hist^{l + 1} \in \allhist$, 
define $\args_{l + 1} = \gbeatsinv(\ar_l) \cap \range(\hist^{l + 1})$; 
define $\dargs[l + 2] = \dbeatsinv(\args_{l + 1})$ as those fragments defending against the anticipated attackers $\args_{l + 1}$; 
and, if $\args_{l + 1} ≠ \emptyset$ (equivalently, if $\dargs[l + 2] ≠ \emptyset$), define $\ar_{l + 2} = \ar_l + \dargs[l + 2]$. 
The sequence presents successive challenges iff, $\forall l \in 2\N \suchthat l ≤ k$, we have that $\ar_l \in \histend[l + 1]$, $\hist^{l + 1} \nileadstost \phi$, and $[\args_{l + 1} = \emptyset ⇔ l = k]$.

@@

A path $\histp$ starts with a path $\hist$ of length $l$ iff $\histp$ has length at least $l$ and $\forall k \in \intvl{1, l}: \histp_k = \hist_k$.
An increasing sequence of paths is a finite sequence $({\hist}^k)_{k \in \intvl{1, N}}$, for some $N \in \N^*$, of paths, 
such that $\forall k \in \intvl{2, N}, {\hist}^k$ starts with ${\hist}^{k-1}$.

Given two incompatible empirical theories for a given normative theory, $\alpha = (\gleadsto[\alpha], \gbeats[\alpha], \alldargs_\alpha, \dbeats[\alpha], +_\alpha)$ and $\beta = (\gleadsto[\beta], \gbeats[\beta], \alldargs_\beta, \dbeats[\beta], +_\beta)$ and given any $i \in I$, the following procedure falsifies one of these theories, queries only $i$, and queries $i$ only about some increasing sequence of paths.
\begin{procedure}
	Pick any $\prop \in \allprops$ about which $\alpha$ and $\beta$ disagree.

@@

Define $C$ a set of criteria, functions $c: \allalts → \R$, each function measuring the performances of the alternatives according to some point of view. Let $t \in \R^C$ be a set of thresholds, thus, consisting of one real valued threshold $t_c \in \R$ per criterion. Say that $a \in \allalts$ is satisfying iff $\forall c \in C: c(a) ≥ t_c$.
 Let $A$ denote the satisfying alternatives.

The theory $\gamma$ claims, intuitively, that individuals $i \in I$ are satisficers using some defined criteria $C$ and thresholds $t = (t_c)_{c \in C}$. Define $\peq$ as $a \peq b$ iff $a \in A \lor b \notin A$. Note that $\peq$ is complete and contains at most two equivalence classes: $A$, and $\allalts \setminus A$. This defines the set of claims that $\gamma$ supports, namely, for any $\emptyset ≠ B \subseteq \allalts$, $\gprops(B) = B \cap A$.

Given $a \in A$, define the fragment $\ar_a$ as the text: “$a$ has performance $c_1(a)$ on criterion $c_1$, and it is reasonable to consider that having at least $t_{c_1}$ is satisfying enough from the point of view captured by $c_1$, [… same phrase for the other criteria…], therefore, it is a fully satisfactory alternative.”
Given $b \in \allalts \setminus A$, define $\ar_b$ as the text: “$b$ is not satisfactory on the criterion [some criterion such that $c(b) < t_{c}$]: the performance $t_c$ should be attained for considering $b$ as satisfying on that point of view, but it obtains only $c(b)$ on that criterion [… same phrase for the other criteria such that $c(b) < t_c$].”

Given a bundle $B$, define $\ar_B$ as the text obtained by concatenating all fragments in $\set{\ar_a \suchthat a \in B}$.

@@

A more realistic variant of this empirical theory would refrain from claiming that all satisfactory alternatives are equivalent, and that all non-satisfactory alternatives are equivalent. Indeed, this claims is likely to fail reflecting the deliberated preferences of the individuals: $i$ will probably pick $a$ rather than $\set{a, b}$ given a bundle of two satisfactory alternatives $\set{a, b}$ with $a$ Pareto-dominating $b$.

To this effect, given $a, b \in \allalts$, define $a \peq b$ iff $a \in A \land b \notin A$ (this also changes the set of claims that $\gamma$ supports) and keep the other definitions intact.
The resulting theory does not claim to capture complete deliberated preferences, as it ignores how any two alternatives that are both satisfactory, or both unsatisfactory, compare. However, whenever there is exactly one satisficing alternative in $B$, it still determines a single deliberated preferred choice for $i$ among $B$.

@@

Given $i \in I, \prop \in \allprops, \ar \in \allargs$ and $\args \subseteq \gbeatsinv(\ar)$, define $\args \ibeats \ar$ iff $\exists \hist \in \allhist \suchthat \args = \gbeatsinv(\ar) \cap \range(\hist) \land \ar \in \histend \land \hist \nileadstost \prop$. 
Such a set $\args$, when not empty, represents an effective set of anticipated attackers of $\ar$ against $\prop$. 
The notation uses a striked through $\prop$ in superscript to suggest that when $\args \ibeats \ar$, the coalition $\args$ argues against $\prop$, from the point of view of $i$.
I write that $\ar$ \emph{$i$-defends} $\prop$ to mean that $\emptyset \nibeats \ar$.

@@

\begin{lemma}
	\label{th:instseq}
	An empirical theory has a falsification instance iff it has a falsification sequence.
\end{lemma}
\begin{proof}
	First, from right to left, consider a falsification sequence $(\hist^{l + 1})_{l \in 2\N, l ≤ k}$ for some $k \in 2\N$, define, as in \cref{sec:fals}, $\args_{l + 1} = \gbeatsinv(\ar_l) \cap \range(\hist^{l + 1})$, $\dargs[l + 2] = \dbeatsinv(\args_{l + 1})$ and (for $l ≤ k - 2$) $\ar_{l + 2} = \ar_l + \dargs[l + 2]$, and let us show that $(i, \phi, \hist^{k + 1}, \ar_{k})$ is a falsification instance. This amounts to show that $\ar_k \in \gargs$, which can be done by observing that $\ar_0 \in \gargs$, and showing that, $\forall l \in 2\N, l ≤ k - 2$, $\ar_l \in \gargs ⇒ \ar_{l + 2} \in \gargs$.
	Indeed, \cref{th:dpiconstr} applies with $\ar_l$ for $\ar_0$ and $\ar_{l + 2}$ for $\ar_2$. As the second member of the equivalence is true by definition of a falsification sequence, $\ar_{l + 2} \in (\dip)(\ar_l)$, and because $\ar_l \in \gargs$, $\ar_{l + 2} \in \gargs$.

	Second, from left to right, consider a falsification instance $(i, \phi, \hist^{k + 1}, \ar_{k})$. As $\ar_k \in \gargs$, $\ar_k \in (\dip)^k({\gleadstoinv}(\prop))$ (equalizing the indices $k$ wlog). Given $\ar_{l + 2} \in (\dip)^j({\gleadstoinv}(\prop))$, with $j > 0$, thus $\ar_{l + 2} \in \dip(\ar_l)$ for some $\ar_l \in (\dip)^{j - 1}({\gleadstoinv}(\prop))$, apply \cref{th:dpiconstr}, obtain $\hist_{l + 1}$. Iterate while $j > 0$. When $j = 0$, the $\hist$ found corresponds to an initial argument.

@@

\begin{theorem}
	\label{th:valid}
	An empirical theory that is normatively adequate and convincing is valid.
\end{theorem}
\begin{proof}[of \cref{th:valid}]
	Given any $i \in I$ and $\prop \in \gpropse$, suffices to show that for some $\ar \in \allargs$, $\ar \ileadstost \prop$; the conclusion then follows thanks to normative adequacy.

@@

>>>>

This article defines decision theories, called theories of \ac{DJ}, that are both normative and descriptive: they claim to be able to yield valid prescriptions, and these prescriptions are claimed to predict reflective behavior. Reflective behavior as defined in this article is observable, by means of observing the reaction of individuals to arguments. This allows to confront theories of \ac{DJ} to empirical data.

@@

Here is an example opposing two simplistic theories of \ac{DJ}, meant to give a flavor of the sort of theories this article aims at defining.

\begin{example}[Two theories about two alternatives]
	Consider a set $\alts = \set{\alt_1, \alt_2}$ of two alternatives whose evaluations on two criteria (say, price and speed, for concreteness), are $(g_1(\alt_1), g_2(\alt_1))$ and $(g_1(\alt_2), g_2(\alt_2))$, respectively, where $g_i: \alts → \R$ are functions giving the performance of alternatives on the criterion $i$.

	Two theories of \ac{DJ} are opposed on their prediction of the reflexive choice of a given set of individuals $I$ having to choose one alternative among $\alts$.
	The \emph{satisfycing} theory claims that any individual $i \in I$ ought to choose, and will reflectively choose, whichever alternative is satisfying (and be indifferent if they are both satisfying or if none of them is). An alternative $\alt$ is satisfying iff $g_1(\alt) < t_1 \land g_2(\alt) > t_2$, where $t_1$ and $t_2$ are two real numbers chosen by the theorist.
	The \emph{additive} theory claims that any individual $i \in I$ ought to choose, and will reflectively choose, whichever alternative $\alt$ maximises $v(\alt) = v_1(g_1(\alt)) + v_2(g_2(\alt))$, where $v_i: \R → \R$ are functions  chosen by the theorist.
	How the parameters of these theories, $t_i$ and $v_i$, are determined is irrelevant to the questions tackled in this article, because it aims at defining and studying theories of \ac{DJ} in general, not at proposing a way of building a specific theory of \ac{DJ}. %For this example, one can assume that these parameters are determined on the basis of observation of the behavior of some population in some similar circumstances.

	Assume that the theories disagree, thus, that $\alt_1$ is satisfying and $\alt_2$ is not (say, it fails on the speed criterion), but $v(\alt_2) > v(\alt_1)$. To determine which theory is right, it is insufficient to simply observe what an individual $i \in I$ effectively chooses. This would merely reflect her unreflective behavior. The approach adopted in this article to capture \acp{DJ} is to declare that the winning theory (if any) is the one that exhibit arguments able to systematically convince individuals that their proposed choice is to be adopted.
	The satisfycing theory could argue as follows: “$\alt_1$ is better than $\alt_2$, because it has a satisfactory price (lower than $t_1$, an acceptable price) and a satisfactory speed (higher than $t_2$, thus an acceptable performance); whereas $\alt_2$ is not satisfactory on the speed criterion”. 
	The additive theory could use the following argument: “$\alt_2$ is better than $\alt_1$, because $\alt_2$ is cheaper than $\alt_1$, and this more than compensates for its weaker performance in terms of speed”.

	Determining which theory is right requires to present both arguments to $i$, then let her choose one alternative in $\alts$.
\end{example}
%The proposed approach starts with this intuition that the reaction of individuals to arguments should matter, and presents a general solution to the problem of capturing reflective judgments or choices. Contrary to the simplistic situation of \cref{ex:two}, this solution must allow to confront more than just two arguments or two theories. It should define precisely what is required for a judgment to be called deliberated, and determine formally which hypothesis are required for the prescriptions of a theory of \ac{DJ} to be valid.

This article is structured as follows.
\Cref{sec:approaches} defines more precisely what is meant here with normative and descriptive theories. These definitions permit to readily see why the apparent impossibility of a claim being both normative and descriptive is an illusion, thereby enabling a solution to what could be called the “normative against descriptive” gap. It also reviews related works in the literature, and briefly sketches the current proposal.
\Cref{sec:normative,sec:descriptive} define the theories of \ac{DJ} and prove that these theories are normative and descriptive in the sense defined in \cref{sec:approaches}. \Cref{sec:preferences} define theories of \ac{DP} on the basis of theories of \ac{DJ}, and \cref{sec:conclusion} concludes.

@@

This gap can thus be overcome by considering reflective descriptive theories, instead of shallow descriptive theories. This is the path this article adopts.

@@

\section{Descriptive theories}
\label{sec:descriptive}
\subsection{Informal presentation}
\commentOCf{Je ne me suis pas encore décidé : descriptive theory or empirical theory ? Pour le moment, j’utilise les deux expressions de façon interchangeable.}
A normative theory defines when a given argument is decisive for a given proposition and a given individual: whenever every sequence of arguments having that argument towards the end of the sequence lead the individual to adhere to the proposition. This, in turn, under condition of accepting its normative validity \commentYM{mais c'est quoi cette "normative validiy" ? de quelles conditions parles-tu ?}, permits in principle to know something about the \ac{DJ} of an individual. But this gives no concrete means to know someone’s \ac{DJ}. The set of finite sequence of arguments is infinite. The definition of decisiveness could admittedly be modified to account for an a priori bound on the maximal length of sequences, but this would not solve the practical problem that this set would be too large to be explored thoroughly 
\commentYM{certain que ce terme existe ?}
\commentOCf{Non, thoroughfully n’existe pas, en effet}. 
Furthermore, only some sequences of arguments can be tested for a given individual: the individual’s patience is limited. A supplementary problem might happen in some situations. Once a sequence has been presented to the individual, this cannot necessarily be undone: she might remember some of the arguments, and they may have a lasting influence on her perspective. Let me call this possibility the hypothesis of memory.
The converse hypothesis amounts to postulating that the individual forgets after some time and use an observation protocol that separates the tests in time sufficiently.
\commentYM{je comprends ce que tu veux dire ici car ça fait un bout de temps qu'on parle de ces question. Je doute cependant qu'un lecteur innocent comprenne un traitre mot}

Achieving knowledge of someone’s \ac{DJ} in practice is tackled by introducing empirical theories of \ac{DJ}. An empirical theory for a given normative theory defines a subset of propositions of the agenda, that it claims are contained in the \acp{DJ} of individuals concerned by the normative theory. They are called the proposition that the theory supports. \commentYM{dans la mesure où, à ce stade, je ne comprends pas bien ce que contient une "normative theory", la séparation avec l' "empirical theory" n'est pas claire pour moi} 
The theory also claims that it is able to argue in favor of the proposition it supports in such a way that individuals will be convinced of the quality of these arguments. This is accomplished by three binary relations, part of the definition of empiricial theories: one represents its initial arguments, one plans the possible attacks, and one defines its counter-arguments. An empirical theory also defines a concatenation operation able to dynamically build arguments from fragments of arguments.
The theory claims that when one of its arguments in support of one of its claims is unconvincing to an individual, it is because of one of the anticipated attacks; and the theory claims to be able to provide counter-arguments to such attacks as they come, that will in turn be either convincing to the individual, or fail to be because of some anticipated attack. The theory is also constrained to have “finite depth”, meaning that it must set an a priori limit to the number of such iterations of attacks and counter-arguments.

Thanks to these operators, an empirical theory may argue in a way that depends on the counter-arguments each individual finds appealing
\commentYM{je ne comprends pas cette phrase, et en particulier le sens qu'il faut donner au terme "adjust"}. Instead of having to give a single decisive argument in support of its claims, the theory may react to exactly the counter-arguments that a given individual thinks relevant.

Such theories satisfy the two properties that are mandatory to achieve what has been announced, namely, to obtain theories of \ac{DJ} whose power rest on a weak normative axiom and on empirically falsifiable claims. First, the main claim of the theory, of being able to argue convincingly, holds if and only if no falsification sequence exist. Second, granted normative adequacy, if indeed a theory is convincing, then it is valid, meaning that the proposition that it claims are \acp{DJ} are indeed \acp{DJ}. These two claims are laid out formally in \cref{th:svalid}.

A second goal of this section is to prove that when two theories disagree about incompatible propositions, one of the theories at least must fail some given series of empirical tests, which the proof describes. Importantly, these tests can be performed on a single individual and are robust against the hypothesis of memory.

@@

\subsection{Validity}
\label{sec:validity}
Given $\args \subseteq \allargs$ and $\hist \in \allhist$,
the set $\args \cap \range(\hist)$ denotes the arguments that are both in $\args$ and contained in the sequence $\hist$.

Define a falsification sequence for $i, \ar_0, \phi$, with $\ar_0 \gleadsto \phi$, thus $\ar_0$ an initial argument for $\phi$, as a finite sequence $\hist^1, \hist^3, …, \hist^{N + 1}$, for some $N \in 2\N$, such that it presents successive challenges, which means the following. 
Given $k \in 2\N \suchthat k ≤ N$, $\ar_k \in \allargs$ and $\hist^{k + 1} \in \allhist$, 
define $\args_{k + 1} = \gbeatsinv(\ar_k) \cap \range(\hist^{k + 1})$; 
define $\dargs[k + 2] = \dbeatsinv(\args_{k + 1})$ as those fragments defending against the anticipated attackers $\args_{k + 1}$; 
and, if $\args_{k + 1} ≠ \emptyset$ (equivalently, if $\dargs[k + 2] ≠ \emptyset$), define $\ar_{k + 2} = \ar_k + \dargs[k + 2]$. 
The sequence presents successive challenges iff, $\forall k \in 2\N \suchthat k ≤ N$, we have that $\ar_k \in \histend[k + 1]$, $\hist^{k + 1} \nileadstost \phi$, and $[\args_{k + 1} = \emptyset ⇔ k = N]$.

@@

A path $\histp$ starts with a path $\hist$ of length $l$ iff $\histp$ has length at least $l$ and $\forall k \in \intvl{1, l}: \histp_k = \hist_k$.
An increasing sequence of paths is a finite sequence $({\hist}^k)_{k \in \intvl{1, N}}$, for some $N \in \N$, of paths, 
such that $\forall k \in \intvl{2, N}, {\hist}^k$ starts with ${\hist}^{k-1}$.

\begin{theorem}
	\label{th:proc}
	Given two incompatible empirical theories for a given normative theory, $\alpha = (\gleadsto[\alpha], \gbeats[\alpha], \alldargs_\alpha, \dbeats[\alpha], +_\alpha)$ and $\beta = (\gleadsto[\beta], \gbeats[\beta], \alldargs_\beta, \dbeats[\beta], +_\beta)$ and given any $i \in I$, there exists a procedure that falsifies one of these theories, queries only $i$, and queries $i$ only about some increasing sequence of paths.
\end{theorem}
That procedure is described in \cref{sec:proc}.

@@

Define $C$ a set of criteria, functions $c: \allalts → \R$, each function measuring the performances of the alternatives according to some point of view. Let $t \in \R^C$ be a set of thresholds, thus, consisting of one real valued threshold $t_c \in \R$ per criterion. Say that $\alt \in \allalts$ is satisfying iff $\forall c \in C: c(\alt) ≥ t_c$.
 Let $\alts$ denote the satisfying alternatives.

The theory $\gamma$ claims, intuitively, that individuals $i \in I$ are satisficers using some defined criteria $C$ and thresholds $t = (t_c)_{c \in C}$. Define $\peq$ as $\alt \peq \altp$ iff $\alt \in \alts \lor \altp \notin \alts$. Note that $\peq$ is complete and contains at most two equivalence classes: $\alts$, and $\allalts \setminus \alts$. This defines the set of claims that $\gamma$ supports, namely, for any $\emptyset ≠ B \subseteq \allalts$, $\gprops(B) = B \cap \alts$.

Given $\alt \in \alts$, define the fragment $\ar_\alt$ as the text: “$\alt$ has performance $c_1(\alt)$ on criterion $c_1$, and it is reasonable to consider that having at least $t_{c_1}$ is satisfying enough from the point of view captured by $c_1$, [… same phrase for the other criteria…], therefore, it is a fully satisfactory alternative.”
Given $\altp \in \allalts \setminus \alts$, define $\ar_\altp$ as the text: “$\altp$ is not satisfactory on the criterion [some criterion such that $c(\altp) < t_{c}$]: the performance $t_c$ should be attained for considering $\altp$ as satisfying on that point of view, but it obtains only $c(\altp)$ on that criterion [… same phrase for the other criteria such that $c(\altp) < t_c$].”

Given a bundle $B$, define $\ar_B$ as the text obtained by concatenating all fragments in $\set{\ar_\alt \suchthat \alt \in B}$.

@@

A more realistic variant of this empirical theory would refrain from claiming that all satisfactory alternatives are equivalent, and that all non-satisfactory alternatives are equivalent. Indeed, this claims is likely to fail reflecting the deliberated preferences of the individuals: $i$ will probably pick $\alt$ rather than $\set{\alt, \altp}$ given a bundle of two satisfactory alternatives $\set{\alt, \altp}$ with $\alt$ Pareto-dominating $\altp$.

To this effect, given $\alt, \altp \in \allalts$, define $\alt \peq \altp$ iff $\alt \in \alts \land \altp \notin \alts$ (this also changes the set of claims that $\gamma$ supports) and keep the other definitions intact.
The resulting theory does not claim to capture complete deliberated preferences, as it ignores how any two alternatives that are both satisfactory, or both unsatisfactory, compare. However, whenever there is exactly one satisficing alternative in $B$, it still determines a single deliberated preferred choice for $i$ among $B$.

@@

Given $i \in I, \prop \in \allprops, \ar \in \allargs$ and $\args \subseteq \gbeatsinv(\ar)$, define $\args \ibeats \ar$ iff $\exists \hist \in \allhist \suchthat \args = \gbeatsinv(\ar) \cap \range(\hist) \land \ar \in \histend \land \hist \nileadstost \prop$. 
Such a set $\args$, when not empty, represents an effective set of anticipated attackers of $\ar$ against $\prop$. 
The notation uses a striked through $\prop$ in superscript to suggest that when $\args \ibeats \ar$, the coalition $\args$ argues successfully against $\prop$ despite the presence of $\ar$, from the point of view of $i$.
I write that $\ar$ \emph{$i$-defends} $\prop$ to mean that $\emptyset \nibeats \ar$.

@@

\begin{lemma}
	\label{th:instseq}
	An empirical theory has a falsification instance iff it has a falsification sequence.
\end{lemma}
\begin{proof}
	First, from right to left, consider a falsification sequence $(\hist^{k + 1})_{k \in 2\N, k ≤ N}$ for some $N \in 2\N$, define, as in \cref{sec:validity}, $\args_{k + 1} = \gbeatsinv(\ar_k) \cap \range(\hist^{k + 1})$, $\dargs[k + 2] = \dbeatsinv(\args_{k + 1})$ and (for $k ≤ N - 2$) $\ar_{k + 2} = \ar_k + \dargs[k + 2]$, and let us show that $(i, \phi, \hist^{N + 1}, \ar_N)$ is a falsification instance. This amounts to show that $\ar_N \in \gargs$, which can be done by observing that $\ar_0 \in \gargs$, and showing that, $\forall k \in 2\N, k ≤ N - 2$, $\ar_k \in \gargs ⇒ \ar_{k + 2} \in \gargs$.
	Indeed, \cref{th:dpiconstr} applies with $\ar_k$ for $\ar_0$ and $\ar_{k + 2}$ for $\ar_2$. As the second member of the equivalence is true by definition of a falsification sequence, $\ar_{k + 2} \in (\dip)(\ar_k)$, and because $\ar_k \in \gargs$, $\ar_{k + 2} \in \gargs$.

	Second, from left to right, consider a falsification instance $(i, \phi, \hist^{N + 1}, \ar_N)$. As $\ar_N \in \gargs$, $\ar_N \in (\dip)^N({\gleadstoinv}(\prop))$ (equalizing the indices $N$ wlog). Given $\ar_{k + 2} \in (\dip)^j({\gleadstoinv}(\prop))$, with $j > 0$, thus $\ar_{k + 2} \in \dip(\ar_k)$ for some $\ar_k \in (\dip)^{j - 1}({\gleadstoinv}(\prop))$, apply \cref{th:dpiconstr}, obtain $\hist_{k + 1}$. Iterate while $j > 0$. When $j = 0$, the $\hist$ found corresponds to an initial argument.

@@

\begin{theorem}
	\label{th:valid}
	An empirical theory that is normatively adequate and convincing is valid.
\end{theorem}
\begin{proof}
	Given any $i \in I$ and $\prop \in \gpropse$, suffices to show that for some $\ar \in \allargs$, $\ar \ileadstost \prop$; the conclusion then follows thanks to normative adequacy.

@@

\section{Comparing theories}
\label{sec:proc}
This section proves \cref{th:proc} constructively, by describing the sequence of tests that must be performed and showing that they must lead to falsifying one of the theories.
To do this, the notion of challenge must be defined and a few simple lemmas are required.

\begin{procedure}
	\label{proc}
	Pick any $\prop \in \allprops$ about which $\alpha$ and $\beta$ disagree.
